---
title: "Rで作図"
subtitle: "1. Introduction"
output:
  ioslides_presentation:
    css: ../css/slide.css
    widescreen: true
date: "2024-03-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

## Contact R server
+ Input 133.45.178.25:8787
+ Fill your ID and PW accordingly
    - ID: your first name
    - common PW: hogefuga1192+

## Contact from outside via VPN
+ Set up your [VPN](https://www.cc.nagasaki-u.ac.jp/service/outside_the_university.html) app such as Cisco AnyConnect
+ https://www.cc.nagasaki-u.ac.jp/service/outside_the_university.html
+ Before connecting the R server, start the VPN

## ID /PW
+ ID: your first name in small letter all
+ PW: your first name in small letter all
+ Beware of
    - CAPITAL / small letter
    - typo

## Project, file, and expansion
+ Project
    - A place (folder) to save all related to a certain analysis
+ file
    - A file to save a R code.
    - Its substance is a .txt file
+ expansion
    - a string to distinguish file type.
    - In R, use .r after a file name. Eg.) hoge.r

## Frequently-used keybinds of the RStudio | NO using mouse for quick operation
+ select ($\uparrow$$\downarrow$$\rightarrow$$\leftarrow$+Shift) $\rightarrow$ Ctrl+Enter: Run the code
+ select $\rightarrow$ Ctrl+Shift+c: Comment out the code
+ Ctrl+Shift+m: %>% (pipe operator; assign an object from right to the first argument of function placed at right)
+ Alt+-: <- (assign operator; assign an object from right to left)
+ **Ctrl+s: save**

## Lesson | NO using Japanese character
1. Make a project entitled "rgraphics.rproj"
2. Make a R file entitled "rgraphics.r"
3. Write your code as follows:
4. Save the R file
5. Show your results

```{r code.sample, include = TRUE, echo=TRUE, message = FALSE}
library(sf)
library(tidyverse)
```

# Fundamental operation

## Where to write your code
1. Source pane (topleft, Crtrl+1)
2. Console pane (Bottomleft, Ctrl+2)
3. Environment pane (Topright\)
4. Output pane (Bottomright)

## Read library
+ Whenever you start the R, read the library first.
+ Library is a mass of functions for specific purposes.
    - R is a station integrating libraries.

```{r load.library, include = TRUE, echo=TRUE}
library(sf)
library(tidyverse)
```

## NOTE
+ Once you install a library, you need not to reinstall it.
+ Once you load a library, you need not to load it again.

```{r note.load.library, include = TRUE, echo=TRUE}
# If necessary, comment out the target code.
# install.packages("tidyverse")
library(tidyverse)
```

## Calculation
+ We can use the R as a calculator
```{r calculator, include = TRUE, echo=TRUE}
1+2+3+4+5+6+7+8+9+10
sum(1:10)

```

## Calculation (Continued)
+ We can use the R as a calculator
```{r calculator.product, include = TRUE, echo=TRUE}
1*2*3*4*5*6*7*8*9*10
prod(1:10)

```

## Calculation (Continued)
+ We can use the R as a calculator
```{r calculator.log, include = TRUE, echo=TRUE}
log(1*2*3*4*5*6*7*8*9*10)
log10(prod(1:10))
```


## Read data
+ R can read variety of files including *rds*, *csv*, *xlsx*, *dta*, and others.
+ This time, we read an MSExcel file.
+ Procedure
    1. Obtain a certain target data
    2. Upload the MSExcel file (.xlsx) onto your project folder.
    3. Read the data following the code.

## Show the data
+ Run the assigned object
+ Frequently-used variable type
    - **dbl**: Double
    - **int**: Integer
    - **chr**: Character
    - **dttm**: date and time (POSIXt)
    - **fct**: factor
+ Depending on the type, what we can do is subject to change.

## Assignment
+ Assign data, figure, and others into an object
+ Free to name the object by your own
+ Available to succeed

```{r assignment, include = TRUE, echo=TRUE}
logresults <- log(1*2*3*4*5*6*7*8*9*10)
logresults
```

## Assignment
+ Download four .rds data file from my GitHub (https://github.com/yuzuruu/lecture/tree/main/r_workshop)
+ Upload the data onto your RStudio server
+ Read any of the data.


## Read objects
```{r read.data, include = TRUE, echo=TRUE}
junior <- 
  readr::read_rds("seniority_data_tutorial_college.rds")
high <- 
  readr::read_rds("seniority_data_tutorial_high.rds")
college <- 
  readr::read_rds("seniority_data_tutorial_college.rds")
university <- 
  readr::read_rds("seniority_data_tutorial_university.rds")
```

## Frequently-used operators
+ <- : assign an object at right to one at left
  - shortcut: Alt+-
+ %>% : assign left-hand-side object into the first argument of the right-hand-side function
  - shortcut: Ctrl+Shift+m


## Combine the data
```{r combine.data, include = TRUE, echo=TRUE}
data_seniority <- 
  junior %>% 
  # bind the object
  # NOTE
  # Name of variables and the N. of variables should be the same.
  dplyr::bind_rows(high) %>% 
  dplyr::bind_rows(college) %>% 
  dplyr::bind_rows(university) 
```

## filter
```{r filter.data.01, include = TRUE, echo=TRUE}
# filter
data_seniority_filtered <- 
  data_seniority  %>% 
  # make a subset of the data using dplyr::filter() function
  # NOTE
  # Multiple conditions are available.
  dplyr::filter(
    school == "university"
  )
# print content of the object
data_seniority_filtered
```

## Frequently-used operators
+ == :equivalent to
+ != :NOT equivalent to
+ &: and
+ |: or
+ %in%: any of (usually used with vector containing conditions)

## filter (Cont.)
```{r filter.data.02, include = TRUE, echo=TRUE}
# filter with multiple variables
data_seniority_filtered <- 
  data_seniority  %>% 
  dplyr::filter(
    school == "university" & gender == "female"
  )
data_seniority_filtered
```

## filter (Cont.)
```{r filter.data.03, include = TRUE, echo=TRUE}
# filter with multiple conditions
data_seniority_filtered <- 
  data_seniority  %>% 
  dplyr::filter(
    school %in% c("university", "college")
  )
data_seniority_filtered
```

# Draw a figure

## What to do
+ Target
    - *Wage census* of Japan (2010-2019)
    - Make a line plot showing wage transition by length of service and age class
    - A basis of so-called "seniotiry payroll"
+ Procedure
    1. Make a data set
    2. Draw the figure
    3. Check its aesthetics
    
## summarize to obtain aggregated data 
```{r summarise, include = TRUE, echo=TRUE}
data_seniority_summary <- 
  data_seniority %>% 
  dplyr::filter(
    school == "university"
  ) %>% 
  group_by(length_service, age_class) %>% 
  summarise(
    Mean = mean(value, na.rm = TRUE)
  )
data_seniority_summary
```

## plot in the beginning 
```{r initial.plot, include = TRUE, echo=TRUE}
line_seniority_summary <- 
  data_seniority_summary %>% 
  ggplot2::ggplot(
    aes(
      x = length_service,
      y = Mean,
      color = age_class,
      group = age_class
    )
  ) +
  geom_point() +
  geom_line()
line_seniority_summary
```

## plot in the beginning (Cont.) | This figure does not satisfies us.
```{r initial.plot.print, include = TRUE, echo=TRUE, warning=FALSE}
line_seniority_summary
```

