---
title: "Statistical hypothesis testing"
subtitle: "R workshop for E-4 members 4"
author: "Yuzuru Utsunomiya"
date: "`r Sys.time()`"
output: 
  ioslides_presentation:
    css: ../css/slide.css
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include = FALSE}
knitr::read_chunk("../rworkshop_code.R")
```

# Statistical test

## Load libraries first
```{r readlibraryandomitna, include = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
```


## The most familiar method

+ Test is a procedure to distinguish whether observations are sampled by accident or not based on theory of probability.
+ Practically, it is employed to compare some statistics such as central tendencies (mean and median), variance, and others.
+ Controversies to be followed
    - Type I error and Type II error ($\alpha$ and $\beta$)
    - Meaning of $p$ value
    - Effect size and power analysis

## Story: Case of comparing 2 groups' means

1. We'd like to compare *mean* by group: whether does a significant (large / small enough) difference between the means exist?
2. The means vary... 
3. We compute a difference of the means while standardizing with SE (t statistics).
4. Whether the computed statistics is significantly large / small enough (different) at a certain level (normally 5% or below), we decide whether the difference exists.

## True number may be there in somewhere
+ We cannot observe the true value $\rightarrow$ We need to estimate the number from sample.
    - Sample mean of $x$: $m_{x}$ (or $\bar{x}$)
    - Population mean of $x$: $\mu_{x}$
+ Most welcome case: $m_{x} = \mu_{x}$
+ To estimate the $\mu$, we collect samples.

## The observations distribute following a certain distribution...

+ Suppose two groups including some observation; $X$ and $Y$.
+ This time, observations of both group are supposed to follow normal distributions with the same variance $\sigma^{2}$.
    - $X \sim N(m_{x}, \sigma^{2})$
    - $Y \sim N(m_{y}, \sigma^{2})$

## Difference between the group $x$ and $y$
+ We denote the difference between mean of $m_x$ and mean of $m_y$, given

$$
d = m_{x} - m_{y}
$$

## $d$ also distributes following a normal distribution.
+ The $d$ also fluctuates. It follows a normal distribution.
    - $d \sim N(\mu_{y}-\mu_{x}, 2\sigma^{2})$
+ The $d$ is affected by variance ($\sigma^{2}$).  
    - So we'd control effect of variance to estimate the true $d$.

## What's $d$?

```{r two.density, include = TRUE, results = FALSE, warning = FALSE, message = FALSE, echo=FALSE}
```

```{r plot.normal.pair.d, include = TRUE, warning = FALSE, message = FALSE, echo=FALSE, fig.align='center'}
```
    
## Using SE of $d$, we control (standardize) the effect.
+ The standardized $d$ (t statistics) is given

$$
t = \frac{m_{y}-m_{x}}{\sqrt{\frac{\sum_{i=1}^{n}(x_{i}-m_{x})^{2}+\sum_{i=1}^{n}(y_{i}-m_{y})^{2}}{(n-1)+(n-1)}\frac{1}{n}}}
= \frac{m_{y}-m_{x}}{\underbrace{\frac{\sigma}{\sqrt{n}}}_{SE}}
$$

## The $t$ follows $t$ distribution.
+ Interpretation: When the $t$ would be larger / smaller enough, there seems to be significant difference between the group.
    - The $t$ is available to approximate using normal distribution.
    - When the $d$ is above 97.5% or below 2.5% of the distribution, let us regard the difference as significant.

## 95% CI (Coefficient Intervals)
```{r plot.density.95.paint, include = TRUE, warning = FALSE, message = FALSE, echo=TRUE, fig.align='center'}
```

# Test in R

## Procedure
1. Check descriptive statistics and distribution
2. test the difference between the group using **t.test** function.
3. Interpret results of test
```{r t.test, include = FALSE, results = FALSE, warning = FALSE, message = FALSE, echo=TRUE}
```

## Sample data
+ Artificially generated two datasets.
    - $x \sim N(0, 1^{2})$  
    - $y \sim N(1, 1^{2})$
+ N = 100 by group

## Code: Generate datasets | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)

```{r sample.ttest, include = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
x.sample <- rnorm(100, mean = 0, sd =1)
y.sample <- rnorm(100, mean = 1, sd =1)
df.x <- data.frame(sample = x.sample, 
                   group = rep("x", 
                               times = length(x.sample)))
df.y <- data.frame(sample = y.sample, 
                   group = rep("y", 
                               times = length(x.sample)))
df.xy <- as.tibble(rbind(df.x, df.y))  
df.xy$group <- factor(df.xy$group) 
```

## Code: Compute descriptive statistics | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r sample.descriptive.statistics, include = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
xy.summary <- df.xy %>% group_by(group) %>% 
  summarise(N = length(sample), 
            Min. = min(sample), 
            Mean = mean(sample), 
            Median = median(sample), 
            Max. = max(sample), 
            SD = sd(sample), 
            SE = sd(sample)/sqrt(length(sample))
            )
```


## Check descriptive statistics | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r summary.table.ttest, include = TRUE, results = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
```

## Code: Draw a density plot | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r densityplot.ttest, include = TRUE, results = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
density.xy <- df.xy %>% 
  ggplot(aes(x=sample, colour = group, fill = group)) + 
  geom_density(alpha = 0.3) + 
  scale_x_continuous(limits = c(-4,4)) +
  theme_bw() +
  theme(axis.line = element_line(size=1, colour="black"),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        panel.background = element_blank(),
        text=element_text(family="Tahoma")
  )
```

## Check distributions | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r density.plot.ttest, include = TRUE, warning = FALSE, message = FALSE, echo=FALSE, fig.align='center'}
```

## t.test | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r t.test.results, include = TRUE, results = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
```

## Notice
+ When a sample distribution does not follow a normal distribution, considering nonparametric tests may help.
    - **wilcox.test**
+ The significant difference does not prove clinical (practical) difference.
    - For interpretation, consider the results carefully in terms of your specialty.
+ To fix an appropriate sample size, refer to **effect size**.

## Exercise | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
+ Using **sleep** data, do the test by group (group 1 and group 2).
+ Gossen, a founder of t-test, made the data when he implemented a medical efficiency test.
```{r t.test.sleep, include = FALSE, results = FALSE, warning = FALSE, message = FALSE, echo=TRUE}
```


## Check descriptive statistics | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r summary.table.sleep, include = TRUE, results = TRUE, warning = FALSE, message = FALSE, echo=FALSE}
```

## Code: Computing descriptive statistics | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r sleep.ds, include = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
data(sleep)
attach(sleep)
sleep <- sleep %>% as.tibble(.)
sleep.summary <- sleep %>% group_by(group) %>% 
  summarise(N = length(extra),
            Min. = min(extra),
            Mean = mean(extra),
            Median = median(extra),
            Max. = max(extra),
            SD = sd(extra),
            SE = sd(extra)/sqrt(length(sample))
            )

```

## Check distributions | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r density.plot.sleep, include = TRUE, warning = FALSE, message = FALSE, echo=FALSE, fig.align='center'}
```

## Code: Drawing density plots | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r sleep.density, include = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
density.sleep <- sleep %>% 
  ggplot(aes(x=extra, colour = group, fill = group)) + 
  geom_density(alpha = 0.3) + 
  scale_x_continuous(limits = c(-4,10)) +
  theme_bw() +
  theme(axis.line = element_line(size=1, colour="black"),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        panel.background = element_blank(),
        text=element_text(family="Tahoma")
  )
```

## t.test | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r t.test.results.sleep, include = TRUE, results = TRUE, warning = FALSE, message = FALSE, echo=FALSE}
```

## Code: Drawing density plots | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r sleep.ttest.01, include = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
sleep.ttest.results <- sleep %>% 
  t.test(extra ~ group, data = ., 
         var.equal = TRUE, 
         paired = FALSE)
```

## Code: Drawing density plots | Please write the code, save it (Ctrl+s), and run (select->Ctrl+Enter)
```{r sleep.ttest.02, include = TRUE, warning = FALSE, message = FALSE, echo=TRUE}
sleep.ttest.results
```

# Next Regression